Unbalanced Mallows Models for Optimizing Expensive Black-Box Permutation Problems

### 2020-10-27 paper/main.tex
fix sec 5

\begin{figure}[htb]
  \centering%
  \begin{minipage}{0.49\linewidth}\centering
    \texttt{N-t1d100.01}\\[-0.2ex]
    \includegraphics[width=\textwidth]{../img/heatmap_N-t1d100_01}
  \end{minipage}
  \begin{minipage}{0.49\linewidth}\centering
    \texttt{N-t1d50.01}\\[-0.2ex]
 \includegraphics[width=\textwidth]{../img/heatmap_N-t1d150_01}
\end{minipage}\\
\includegraphics[width=0.49\textwidth]{../img/heatmap_rec05}
  \includegraphics[width=0.49\textwidth]{../img/heatmap_kra32_dat}
    \caption{Mean fitness  for different configuration for the learning rate \label{fig:heatmaps}}
\end{figure}

The plot shows that UMM is indeed very sensitive to the values of $r_1$ and
$r_2$, yet, low values of $r_1$ and high values of $r_2$ almost always produce
the best results, independently of the problem. Nevertheless, there are a few
cases where an interaction seems to occur and particular combinations perform
remarkably well.\MANUEL{why?}  In particular, we observe that the pattern on
the two left-most figures differs from the two right-most ones. While on the
left the fitness gets worse homogeneously from the top right corner to the
bottom left one, the pattern on the right is not as homogeneous, yet still
fitness gets worse with decreasing $r_2$.  We observe that this behavior is
related to the adequacy of UMM to the problem: the homogeneous behavior of
parameters $r_1$ and $r_2$ depends on the problem (not to the particular
instances) and the more homogeneous the behavior, the better performance of
UMM. We elaborate this argument in the following sections with detail.

Over all the instances \MANUEL{X (which instances we tested? all synthetic?)} we
see that the configuration $r_1 =0.1$, $r_2=0.9$ has good
performance\MANUEL{how do you define this? The best mean?}. From this
preliminary experiment, it is not obvious how the best settings are related to
problem features, such as permutation length ($n$). On the other hand, given
the sharp transitions in the plots, it is clear that further fine-tuning may
improve performance and uncover further patterns.  Although automatic parameter
tuning may likely provide even more fine-tuned settings, our goal here is to
understand how UMM performs in comparison to an existing Bayesian method (CEGO)
and automatically tuning the parameters of CEGO is not computationally
feasible.  For consistency, in the following sections, we run UMM with this
parameter configuration. Moreover, we recommend this configuration as the
current default of UMM.

@@

\EKHINE{cambio un poco al argumento}
When we compare the behaviour of CEGO and UMM on synthetic LOP instances
(Fig.~\ref{fig:lop_synth}), we observe a clear pattern independently of the
uniformity of the instances. In particular, we see a linear convergence to the optimal value for UMM. This is due to the linear decrease in the parameter $\theta$ (line~\ref{line:theta} in Algorithm~\ref{alg:umm}), which imposes a linear variance decrease on the internal probability model of the algorithm over the $m$ functions evaluations. Overall, UMM found the best solution in all the repetitions.

>>>>

\begin{figure}[htb]
  \centering%
  \begin{minipage}{0.49\linewidth}\centering
    \texttt{N-t1d100.01}\\[-0.2ex]
    \includegraphics[width=\textwidth]{../img/heatmap_N-t1d100_01}
  \end{minipage}
  \begin{minipage}{0.49\linewidth}\centering
    \texttt{N-t1d50.01}\\[-0.2ex]
 \includegraphics[width=\textwidth]{../img/heatmap_N-t1d150_01}
\end{minipage}\\
\includegraphics[width=0.49\textwidth]{../img/heatmap_rec05}
  \includegraphics[width=0.49\textwidth]{../img/heatmap_kra32_dat}
    \caption{Mean fitness  for different configuration for the learning rate (darker is better). \label{fig:heatmaps}}
\end{figure}

The plot shows that UMM is indeed very sensitive to the values of $r_1$ and
$r_2$, yet, low values of $r_1$ and high values of $r_2$ almost always produce
the best results, independently of the problem. Nevertheless, there are a few
cases where an interaction seems to occur and particular combinations perform
remarkably well.\MANUEL{Missing data, whitespaces}  In particular, we observe that the pattern on
the two left-most figures differs from the two right-most ones. While on the
left the fitness gets worse homogeneously from the top right corner to the
bottom left one, the pattern on the right is not as homogeneous, yet still
fitness gets worse with decreasing $r_2$.  We observe that this behavior is
related to the adequacy of UMM to the problem: the homogeneous behavior of
parameters $r_1$ and $r_2$ depends on the problem (not to the particular
instances) and the more homogeneous the behavior, the better performance of
UMM. We elaborate this argument in the following sections with detail.

Over all the instances  we
see that the configuration $r_1 =0.1$, $r_2=0.9$ has good
performance\MANUEL{how do you define this? The best mean?}. From this
preliminary experiment, it is not obvious how the best settings are related to
problem features, such as permutation length ($n$). On the other hand, given
the sharp transitions in the plots, it is clear that further fine-tuning may
improve performance and uncover further patterns.  Although automatic parameter
tuning may likely provide even more fine-tuned settings, our goal here is to
understand how UMM performs in comparison to an existing Bayesian method (CEGO)
and automatically tuning the parameters of CEGO is not computationally
feasible.  For consistency, in the following sections, we run UMM with this
parameter configuration. Moreover, we recommend this configuration as the
current default of UMM.

@@

When we compare the behaviour of CEGO and
UMM on synthetic LOP instances (Fig.~\ref{fig:lop_synth}), we observe a clear
pattern independently of the uniformity of the instances. In particular, we see
a linear convergence to the optimal value for UMM.  This is due to the linear
decrease in the parameter $\theta$ (line~\ref{line:theta} in
Algorithm~\ref{alg:umm}), which imposes a linear variance decrease on the
internal probability model of the algorithm over the $m$ functions
evaluations. Overall, UMM found  the best solutions in all the repetitions.\MANUEL{in all runs UMM found a better solution? }
